---
title: Create CMs
---

# Set-up

- Load libraries

```{r, echo=FALSE, message=FALSE}
library(magrittr)
library(data.table)
library(ggplot2)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(knitr)
library(lme4)
library(emmeans)
library(binhf)
```

- Set knitr to ommit code chunks

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

- Set up file paths
```{r}
base_dir = file.path(path.expand('~'),
                     'direction_decoding',
                     fsep=.Platform$file.sep)
data_dir = file.path(base_dir, 'decoding',
                     'train_beta_test_beta',
                     'results',
                     fsep = .Platform$file.sep)
save_dir= file.path(base_dir,
                    'analysis',
                    'curve_fitting',
                    fsep = .Platform$file.sep)
highres_dir = file.path(base_dir,
                        'decoding',
                        'train_beta_test_raw',
                        'results', fsep = .Platform$file.sep)
```


- Set up list of masks and subjects

```{r}
# List of masks
masks = c('rsc', 'v1')

# Get list of subjects from classifier accuracy
file = file.path(data_dir,'clf_accuracy.tsv', fsep=.Platform$file.sep)
sub_list = unique(fread(file, select=1, sep='\t'))

# Get index of old and young participants
old_idx = grepl('old', sub_list$sub_id)
young_idx = grepl('young', sub_list$sub_id)
```

---

# Get fitted curves of Gaussian and Uniform

- Fitting functions

```{r}
# Define Gauss function 
Mygauss = function(params, angle) {
	m = 0
	sd = params[1]
	fit = exp(-0.5 * ((angle - m)*sd)^2)
	fit = fit/sum(fit)
	return(fit)
}

# Define uniform function
Myuniform = function(params, angle) {
	peak = params[1]
	fit = rep((1-peak)/5, 6)
	fit[3] = peak
	return(fit)
}

# Cost function, here likelihood assuming arbitrary SD of 0.25 of normally distributed error term 
Mycost = function(fit, data) {
	se = (data - fit)^2
	sse = sum(se)
	like = 1/sqrt(2*pi*0.25^2) * exp(-se/(2*0.25^2))
	LL = -sum(log(like))
	return(LL)
}

# Wrapper for Gauss cost
Mygauss_sse = function(params, angle, data, cidx) {
	fit = Mygauss(params, angle)
	LL = Mycost(fit[cidx], data[cidx])
}

# Wrapper for uniform cost
Myuniform_sse = function(params, angle, data, cidx) {
	fit = Myuniform(params, angle)
	LL = Mycost(fit[cidx], data[cidx])
	return(LL) 
}

# Function to get SSE
Mysse = function(fit, data) {
  se = (data - fit)^2
  sse = sum(se)
  return(sse)
}
```

- Read prediction of beta decoding

```{r}
path = file.path(data_dir, 'clf_prediction.tsv', fsep=.Platform$file.sep)
df_prediction = read.table(path, sep='\t', header = TRUE)
```

- Add group column
```{r}
df_prediction$group = NA
df_prediction$group[grep('old', df_prediction$sub_id)] = 'older'
df_prediction$group[grep('young', df_prediction$sub_id)] = 'younger'
```

- Set bins of beta decoding (sic equally sized direction bins)

```{r}
bins = unique(df_prediction$angle_bin)
n_bins = length(bins)
```

- Make CM direction invariant

```{r, echo=FALSE}
df_prediction_invariant_subj = data.frame(matrix(NA,
                                            0,
                                            length(unique(df_prediction$angle_bin))+2))
colnames(df_prediction_invariant_subj) = as.character(
  c(
    c(seq(length(bins))),
    'mask',
    'sub_id'))

# Loop over each mask
for(mask_count in c(1:length(masks))){
  
  # Select mask
  mask = masks[mask_count]
  
  for(sub_count in 1:length(sub_list$sub_id)){
    
    # Loop over each participant
    sub_id = sub_list[sub_count]$sub_id
    
    # Allocate matrix to hold bin specific counts
    counts = data.frame(matrix(NA, length(bins), ncol(df_prediction_invariant_subj)-2))
    
    for(i in unique(df_prediction$prediction)){
      
      # Constrain data to mask
      data = df_prediction[which(df_prediction$mask == mask),]
      # Constrain data to specific prediction
      data = data[which(data$sub_id == sub_id & data$prediction == i),]
      
      # Count number of prediction given a certain traveled direction
      counts[i,] = table(factor(data$angle_bin, levels=c(bins)))
      
      # Shift correct direction to first position
      counts[i,] = shift(counts[i,], places=(i-1), dir='left')
    }
    
    # Collapse over directions
    counts = data.frame(t(colSums(counts)))
    
    # Normalize counts
    counts = counts/sum(counts)
    
    # Shift correct direction to center (for plotting):
    counts = data.frame(t(apply(counts, 1, function(x) shift(x, places = 2,
                                                             dir = 'right'))))
    
    # Add mask variable
    counts$mask = mask
    # Add sub_id variable
    counts$sub_id = sub_id
    
    # Set column names
    colnames(counts) = as.character(c((bins - bins[length(bins)/2]) * 60, 'mask', 'sub_id'))
    
    # Append mask specific value to data frame
    df_prediction_invariant_subj = rbind(df_prediction_invariant_subj, counts)
    
  }
}
```

- Add group variable and reorder columns

```{r}
df_prediction_invariant_subj$group = NA
df_prediction_invariant_subj$group[grep('old', df_prediction_invariant_subj$sub_id)] = 'older'
df_prediction_invariant_subj$group[grep('young', df_prediction_invariant_subj$sub_id)] = 'younger'

df_prediction_invariant_subj = df_prediction_invariant_subj[,c(n_bins + 2,n_bins + 1,n_bins + 3,1:n_bins)]
```

- Change column names to direction invariant

```{r}
colnames(df_prediction_invariant_subj) = (
  c('sub_id',
    'mask',
    'group',
    paste(
      colnames(df_prediction_invariant_subj)[4:length(colnames(df_prediction_invariant_subj))],
      'deg',
      sep='')))
```

- Save CM to file

```{r}
file = file.path(save_dir, 'data_cm.tsv')
write.table(df_prediction_invariant_subj, file, sep='\t', row.names = FALSE)
```


# Fit uniform and gauss to CM and save as files

```{r}
angles = (bins - bins[length(bins)/2]) * 60
fit_cols = c(1:((length(bins)/2)-1),((length(bins)/2)+1):length(bins))

# Fitting functions combining previous functions
Gaussfit <- function(confusion_mat, angles){optim(c(0.01),
                                                  function(params) {Mygauss_sse(params,
                                                                                angles,
                                                                                confusion_mat,
                                                                                fit_cols)},
                                                  method = 'L-BFGS-B',
                                                  lower = 0,
                                                  upper = 1)}
Unifit <- function(confusion_mat, angles){optim(c(0.01),
                                                function(params) {Myuniform_sse(params,
                                                                                angles,
                                                                                confusion_mat,
                                                                                fit_cols)},
                                                method="L-BFGS-B",
                                                lower = 0,
                                                upper = 1)}

# Apply functions to each line of subject wise confusion matrices
# Gauss (Precision, LL of fit, SSE)
df_prediction_invariant_subj$prec = apply(df_prediction_invariant_subj,
                                          1,
                                          function(x) Gaussfit(as.numeric(x[4:(length(bins)+3)]),
                                                               angles)$par)
df_prediction_invariant_subj$ll_gauss = apply(df_prediction_invariant_subj,
                                              1,
                                              function(x) Gaussfit(as.numeric(x[4:(length(bins)+3)]),
                                                                   angles)$value)
df_prediction_invariant_subj$sse_gauss = apply(df_prediction_invariant_subj,
                                               1,
                                               function(x) sum(
                                                 (as.numeric(x[fit_cols+3]) - 
                                                    Mygauss(as.numeric(x[4:(length(bins)+3)]),
                                                            angles)[fit_cols])^2))
# Uniform (LL of fit, SSE)
df_prediction_invariant_subj$ll_uni = apply(df_prediction_invariant_subj,
                                            1,
                                            function(x) Unifit(as.numeric(x[4:(length(bins)+3)]),
                                                               angles)$value)
df_prediction_invariant_subj$sse_uni = apply(df_prediction_invariant_subj,
                                             1,
                                             function(x) sum(
                                               (as.numeric(x[fit_cols+3]) - 
                                                  Myuniform(as.numeric(x[4:(length(bins)+3)]),
                                                            angles)[fit_cols])^2))

```

- Get only fit data and save

```{r}
# Exclude CM from data to get only parameters of fit
data_fit = df_prediction_invariant_subj[,-c(4:9)]

# Save data
file = file.path(save_dir, 'curve_fit.tsv')
write.table(data_fit, file, sep='\t', row.names = FALSE)
```

- Fit Gauss (via precision taken from curve fits)

```{r}
data_fit_gauss = df_prediction_invariant_subj[,c(1:9)]
data_fit_gauss[,4:ncol(data_fit_gauss)] = data.frame(
  t(
    apply(data_fit, 1, function(x) Mygauss(as.numeric(x[4]), angles))))
```

- Fit uniform (via peak of CM)

```{r}
data_fit_uni = df_prediction_invariant_subj[,c(1:9)]
data_fit_uni[,4:ncol(data_fit_gauss)] = data.frame(
  t(
    apply(df_prediction_invariant_subj, 1, function(x) Myuniform(as.numeric(x[6]), angles))))
```

- Save fitted curves to .tsv

```{r}
file = file.path(save_dir, 'data_gauss.tsv')
write.table(data_fit_gauss, file, sep='\t', row.names = FALSE)

file = file.path(save_dir, 'data_uni.tsv')
write.table(data_fit_uni, file, sep='\t', row.names = FALSE)
```

---

# High resolution confusion matrices

- Read prediction of continuous decoding

```{r}
path = file.path(highres_dir, 'clf_prediction.tsv', fsep=.Platform$file.sep)
df_prediction = read.table(path, sep='\t', header = TRUE)
```

- Add group column
```{r}
df_prediction$group = NA
df_prediction$group[grep('old', df_prediction$sub_id)] = 'older'
df_prediction$group[grep('young', df_prediction$sub_id)] = 'younger'
```

- Add column with higher bin resolution

```{r}
# Define bin resolution (possible values: 60/bin_resolution leaves no modulus)
bin_resolution = 10

df_prediction$bin_mode_res = floor(df_prediction$angle_mode/bin_resolution)
df_prediction$bin_mean_res = floor(df_prediction$angle_mean/bin_resolution)

n_bins = 60/bin_resolution * 6

bins = seq(0, 359.99, bin_resolution)
```

- Make CM direction invariant
```{r, echo=FALSE}
df_prediction_invariant_subj = data.frame(matrix(NA,
                                            0,
                                            length(unique(df_prediction$bin_mean_res))+2))
colnames(df_prediction_invariant_subj) = as.character(c(bins - bins[length(bins)/2], 'mask', 'sub_id'))

# Loop over each mask
for(mask_count in c(1:length(unique(df_prediction$mask)))){
  
  # Select mask
  mask = masks[mask_count]
  
  for(sub_count in 1:length(sub_list$sub_id)){
    
    # Loop over each participant
    sub_id = sub_list[sub_count]$sub_id
    
    # Allocate matrix to hold bin specific counts
    counts = data.frame(matrix(NA, 6, ncol(df_prediction_invariant_subj)-2))
    
    for(i in unique(df_prediction$prediction)){
      
      data = df_prediction[which(df_prediction$mask == mask),]
      data = data[which(data$sub_id == sub_id & data$prediction == i),]
      
      # Count number of prediction given a certain traveled direction
      counts[i,] = table(factor(data$bin_mean_res, levels=c(0:(length(bins)-1))))
      
      # Shift correct direction to first position
      counts[i,] = shift(counts[i,], places=((i-1)*60/bin_resolution), dir='left')
    }
    
    # Collapse over directions
    counts = data.frame(t(colSums(counts)))
    
    # Normalize counts
    counts = counts/sum(counts)
    
    # Shift correct direction to center (for plotting): Most middle, left side bin center
    counts = data.frame(t(apply(counts, 1, function(x) shift(x, places=(120/bin_resolution) +
                                                               floor((60/bin_resolution)/2), dir='right'))))
    
    # Add mask variable
    counts$mask = mask
    # Add sub_id variable
    counts$sub_id = sub_id
    
    # Set column names
    colnames(counts) = as.character(c(bins - bins[length(bins)/2], 'mask', 'sub_id'))
    
    # Append mask specific value to data frame
    df_prediction_invariant_subj = rbind(df_prediction_invariant_subj, counts)
    
  }
}
```

- Add group variable and reorder columns

```{r}
df_prediction_invariant_subj$group = NA
df_prediction_invariant_subj$group[grep('old', df_prediction_invariant_subj$sub_id)] = 'older'
df_prediction_invariant_subj$group[grep('young', df_prediction_invariant_subj$sub_id)] = 'younger'

df_prediction_invariant_subj = df_prediction_invariant_subj[,c(n_bins + 2,n_bins + 1,n_bins + 3,1:n_bins)]
```

- Change column names to direction invariant

```{r}
colnames(df_prediction_invariant_subj) = c('sub_id', 'mask', 'group',
                                           paste(c(bins - bins[length(bins)/2]),'deg', sep=''))
```

- Save high res CM to file

```{r}
file = file.path(save_dir, 'data_cm_highres.tsv')
write.table(df_prediction_invariant_subj, file, sep='\t', row.names = FALSE)
```

---

# Fit uniform and gauss to highres and save as files

- Define functions for curve fitting

```{r}
# Define Function for curve fitting
Mygauss = function(params, angle) {
	m = 0
	sd = params[1]
	fit = exp(-0.5 * ((angle - m)*sd)^2)
	fit = fit/sum(fit)
	return(fit)
}

# Define uniform function
Myuniform = function(params, angle) {
	peak = params[1]
	fit = rep((1-peak)/(n_bins-1), n_bins)
	fit[60/bin_resolution * 3] = peak
	return(fit)
}

# Cost function, here likelihood assuming arbitrary SD of 0.25 of normally distributed error term 
Mycost = function(fit, data) {
	se = (data - fit)^2
	sse = sum(se)
	like = 1/sqrt(2*pi*0.25^2) * exp(-se/(2*0.25^2))
	LL = -sum(log(like))
	return(LL)
}

# Wrapper for Gauss cost
Mygauss_sse = function(params, angle, data, cidx) {
	fit = Mygauss(params, angle)
	LL = Mycost(fit[cidx], data[cidx])
}

# Wrapper for uniform cost
Myuniform_sse = function(params, angle, data, cidx) {
	fit = Myuniform(params, angle)
	LL = Mycost(fit[cidx], data[cidx])
	return(LL) 
}
```

- Make CM direction invariant

```{r, echo=FALSE}
df_prediction_invariant_subj = data.frame(matrix(NA,
                                            0,
                                            length(unique(df_prediction$bin_mean_res))+2))
colnames(df_prediction_invariant_subj) = as.character(c(bins - bins[length(bins)/2], 'mask', 'sub_id'))

# Loop over each mask
for(mask_count in c(1:length(masks))){
  
  # Select mask
  mask = masks[mask_count]
  
  for(sub_count in 1:length(sub_list$sub_id)){
    
    # Loop over each participant
    sub_id = sub_list[sub_count]$sub_id
    
    # Allocate matrix to hold bin specific counts
    counts = data.frame(matrix(NA, 6, ncol(df_prediction_invariant_subj)-2))
    
    for(i in unique(df_prediction$prediction)){
      
      data = df_prediction[which(df_prediction$mask == mask),]
      data = data[which(data$sub_id == sub_id & data$prediction == i),]
      
      # Count number of prediction given a certain traveled direction
      counts[i,] = table(factor(data$bin_mean_res, levels=c(0:(length(bins)-1))))
      
      # Shift correct direction to first position
      counts[i,] = shift(counts[i,], places=((i-1)*60/bin_resolution), dir='left')
    }
    
    # Collapse over directions
    counts = data.frame(t(colSums(counts)))
    
    # Normalize counts
    counts = counts/sum(counts)
    
    # Shift correct direction to center (for plotting): Most middle, left side bin center
    counts = data.frame(t(apply(counts, 1, function(x) shift(x, places=(120/bin_resolution) +
                                                               floor((60/bin_resolution)/2), dir='right'))))
    
    # Add mask variable
    counts$mask = mask
    # Add sub_id variable
    counts$sub_id = sub_id
    
    # Set column names
    colnames(counts) = as.character(c(bins - bins[length(bins)/2], 'mask', 'sub_id'))
    
    # Append mask specific value to data frame
    df_prediction_invariant_subj = rbind(df_prediction_invariant_subj, counts)
    
  }
}
```

- Get model parameters to extract curves

```{r}
angles = bins - bins[length(bins)/2]
fit_cols = c(1:((length(bins)/2)-1),((length(bins)/2)+1):length(bins))

# Fitting functions combining previous functions
Gaussfit <- function(confusion_mat, angles){optim(c(0.01),
                                                  function(params) {Mygauss_sse(params,
                                                                                angles,
                                                                                confusion_mat,
                                                                                fit_cols)},
                                                  method = 'L-BFGS-B',
                                                  lower = 0,
                                                  upper = 1)}
Unifit <- function(confusion_mat, angles){optim(c(0.01),
                                                function(params) {Myuniform_sse(params,
                                                                                angles,
                                                                                confusion_mat,
                                                                                fit_cols)},
                                                method="L-BFGS-B",
                                                lower = 0,
                                                upper = 1)}

# Apply functions to each line of subject wise confusion matrices
# Gauss (Precision, LL of fit, SSE)
df_prediction_invariant_subj$prec = apply(df_prediction_invariant_subj,
                                          1,
                                          function(x) Gaussfit(as.numeric(x[1:length(bins)]),
                                                               angles)$par)
df_prediction_invariant_subj$ll_gauss = apply(df_prediction_invariant_subj,
                                              1,
                                              function(x) Gaussfit(as.numeric(x[1:length(bins)]),
                                                                   angles)$value)
df_prediction_invariant_subj$sse_gauss = apply(df_prediction_invariant_subj,
                                               1,
                                               function(x) sum(
                                                 (as.numeric(x[fit_cols]) - 
                                                    Mygauss(as.numeric(x[length(bins)+3]),
                                                            angles)[fit_cols])^2))
# Uniform (LL of fit, SSE)
df_prediction_invariant_subj$ll_uni = apply(df_prediction_invariant_subj,
                                            1,
                                            function(x) Unifit(as.numeric(x[1:length(bins)]),
                                                               angles)$value)
df_prediction_invariant_subj$sse_uni = apply(df_prediction_invariant_subj,
                                             1,
                                             function(x) sum(
                                               (as.numeric(x[fit_cols]) - 
                                                  Myuniform(as.numeric(x[length(bins)/2]),
                                                            angles)[fit_cols])^2))

```

- Extract fitted curves

```{r}
data_fit_gauss = df_prediction_invariant_subj[,1:(length(bins)+2)]
data_fit_uni = df_prediction_invariant_subj[,1:(length(bins)+2)]

# Fill uni fit and gauss fit
data_fit_gauss[,1:length(bins)] = t(apply(df_prediction_invariant_subj,
                              1,
                              function(x) Mygauss(as.numeric(x[length(bins) + 3]),angles)))
data_fit_uni[,1:length(bins)] = t(apply(df_prediction_invariant_subj,
                              1,
                              function(x) Myuniform(as.numeric(x[length(bins)/2]),angles)))
```

- Add group column, sort columns, and rename columns

```{r}
# Add group column
data_fit_gauss$group = NA
data_fit_gauss$group[grep('old', data_fit_gauss$sub_id)] = 'older'
data_fit_gauss$group[grep('young', data_fit_gauss$sub_id)] = 'younger'
# Sort columns
data_fit_gauss = data_fit_gauss[,c(n_bins + 2, n_bins + 1, n_bins + 3, 1:n_bins)]
# Rename columns
colnames(data_fit_gauss) = c('sub_id', 'mask', 'group',
                             paste(c(bins - bins[length(bins)/2]),'deg', sep=''))

# Add group column
data_fit_uni$group = NA
data_fit_uni$group[grep('old', data_fit_uni$sub_id)] = 'older'
data_fit_uni$group[grep('young', data_fit_uni$sub_id)] = 'younger'
# Sort columns
data_fit_uni = data_fit_uni[,c(n_bins + 2, n_bins + 1, n_bins + 3, 1:n_bins)]
# Rename columns
colnames(data_fit_uni) = c('sub_id', 'mask', 'group',
                             paste(c(bins - bins[length(bins)/2]),'deg', sep=''))
```

- Save fitted curves to .tsv

```{r}
file = file.path(save_dir, 'data_gauss_highres.tsv')
write.table(data_fit_gauss, file, sep='\t', row.names = FALSE)

file = file.path(save_dir, 'data_uni_highres.tsv')
write.table(data_fit_uni, file, sep='\t', row.names = FALSE)
```


