---
title: "train_raw_test_raw_plotting"
output:
  html_document:
    df_print: paged
---

# Load libraries
```{r, message=FALSE}
library(reshape2)
library(ggplot2)
library(scales)
library(plotly)
library(plyr)
library(Rfast)
library(data.table)
library(knitr)
library(binhf)
library(viridis)
library(grid)
library(gridExtra)
library(lme4)
library(emmeans)
```


- Set all chunks to no echo
```{r}
knitr::opts_knit$set(echo=FALSE)
```

---
---

# Set up

- Set up paths

```{r, echo=FALSE}
bids_dir = file.path(path.expand('~'),
                     'direction_decoding_BIDS',
                     fsep = .Platform$file.sep)
repo_dir = file.path(path.expand('~'),
                     'direction_decoding',
                     fsep = .Platform$file.sep)
fwd_dir = file.path(repo_dir,
                    'decoding',
                    'train_beta_test_raw',
                    'results',
                    fsep = .Platform$file.sep)
bwd_dir = file.path(repo_dir, 'decoding',
                    'backwards_decoding',
                    'train_beta_test_raw',
                    'results',
                    fsep = .Platform$file.sep)
```

- Read in .tsv-files
    - Classification accuracy and prediction of FWD walking and BWD walking
    
```{r, echo=FALSE}
# Accuracy
# FWD
path = file.path(fwd_dir, 'clf_accuracy.tsv', fsep=.Platform$file.sep)
df_acc_fwd = read.table(path, sep='\t', header = TRUE)
# BWD
path = file.path(bwd_dir, 'clf_accuracy.tsv', fsep=.Platform$file.sep)
df_acc_bwd = read.table(path, sep='\t', header = TRUE)
# Prediction
# FWD
path = file.path(fwd_dir, 'clf_prediction.tsv', fsep=.Platform$file.sep)
df_pred_fwd = read.table(path, sep='\t', header = TRUE)
# BWD
path = file.path(bwd_dir, 'clf_prediction.tsv', fsep=.Platform$file.sep)
df_pred_bwd = read.table(path, sep='\t', header = TRUE)
```

---

- Get list of masks and subjects
```{r, echo=FALSE}
masks = as.vector(unique(df_acc_fwd$mask))
sub_list = as.vector(unique(df_pred_bwd$sub_id))
```

- Add group column to data frames
```{r, echo=FALSE}
# Accuracy
# FWD
df_acc_fwd$group = NA
df_acc_fwd$group[grep('old', df_acc_fwd$sub_id)] = 'older'
df_acc_fwd$group[grep('young', df_acc_fwd$sub_id)] = 'younger'
# BWD
df_acc_bwd$group = NA
df_acc_bwd$group[grep('old', df_acc_bwd$sub_id)] = 'older'
df_acc_bwd$group[grep('young', df_acc_bwd$sub_id)] = 'younger'

# Prediction
# FWD
df_pred_fwd$group = NA
df_pred_fwd$group[grep('old', df_pred_fwd$sub_id)] = 'older'
df_pred_fwd$group[grep('young', df_pred_fwd$sub_id)] = 'younger'
# BWD
df_pred_bwd$group = NA
df_pred_bwd$group[grep('old', df_pred_bwd$sub_id)] = 'older'
df_pred_bwd$group[grep('young', df_pred_bwd$sub_id)] = 'younger'
```

---

# How does the accuracy at the walking direction target change based on FWD and BWD walking in different masks?

- In case there are shifts in the accuracy:
    - To find out if the lost accuracy shifts towards the 180° phase shifted direction (as predicted by 
    classification based on visual features) we need to assess the amount of 180° phase shifted predictions.

- Rebuild accuracy values out of predictions for FWD and BWD files

```{r, echo=FALSE}
# FWD
df_pred_fwd_inv = df_pred_fwd
df_pred_fwd_inv$bin_diff_mean = df_pred_fwd_inv$prediction - df_pred_fwd_inv$bin_mean
df_pred_fwd_inv$bin_diff_mode = df_pred_fwd_inv$prediction - df_pred_fwd_inv$bin_mode

# Mean
# Make prediction direction invariant
# Dir 1
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 1 & df_pred_fwd_inv$bin_diff_mean == 4] = -2
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 1 & df_pred_fwd_inv$bin_diff_mean == 5] = -1
# Dir 2
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 2 & df_pred_fwd_inv$bin_diff_mean == 4] = -2
# Dir 4
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 4 & df_pred_fwd_inv$bin_diff_mean == -3] = 3
# Dir 5
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 5 & df_pred_fwd_inv$bin_diff_mean == -4] = 2
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 5 & df_pred_fwd_inv$bin_diff_mean == -3] = 3
# Dir 6
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 6 & df_pred_fwd_inv$bin_diff_mean == -5] = 1
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 6 & df_pred_fwd_inv$bin_diff_mean == -4] = 2
df_pred_fwd_inv$bin_diff_mean[df_pred_fwd_inv$bin_mean == 6 & df_pred_fwd_inv$bin_diff_mean == -3] = 3

# Mode
# Make prediction direction invariant
# Dir 1
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 1 & df_pred_fwd_inv$bin_diff_mode == 4] = -2
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 1 & df_pred_fwd_inv$bin_diff_mode == 5] = -1
# Dir 2
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 2 & df_pred_fwd_inv$bin_diff_mode == 4] = -2
# Dir 4
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 4 & df_pred_fwd_inv$bin_diff_mode == -3] = 3
# Dir 5
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 5 & df_pred_fwd_inv$bin_diff_mode == -4] = 2
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 5 & df_pred_fwd_inv$bin_diff_mode == -3] = 3
# Dir 6
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 6 & df_pred_fwd_inv$bin_diff_mode == -5] = 1
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 6 & df_pred_fwd_inv$bin_diff_mode == -4] = 2
df_pred_fwd_inv$bin_diff_mode[df_pred_fwd_inv$bin_mode == 6 & df_pred_fwd_inv$bin_diff_mode == -3] = 3



# BWD
df_pred_bwd_inv = df_pred_bwd
df_pred_bwd_inv$bin_diff_mean = df_pred_bwd_inv$prediction - df_pred_bwd_inv$bin_mean
df_pred_bwd_inv$bin_diff_mode = df_pred_bwd_inv$prediction - df_pred_bwd_inv$bin_mode

# Mean
# Make prediction direction invariant
# Dir 1
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 1 & df_pred_bwd_inv$bin_diff_mean == 4] = -2
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 1 & df_pred_bwd_inv$bin_diff_mean == 5] = -1
# Dir 2
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 2 & df_pred_bwd_inv$bin_diff_mean == 4] = -2
# Dir 4
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 4 & df_pred_bwd_inv$bin_diff_mean == -3] = 3
# Dir 5
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 5 & df_pred_bwd_inv$bin_diff_mean == -4] = 2
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 5 & df_pred_bwd_inv$bin_diff_mean == -3] = 3
# Dir 6
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 6 & df_pred_bwd_inv$bin_diff_mean == -5] = 1
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 6 & df_pred_bwd_inv$bin_diff_mean == -4] = 2
df_pred_bwd_inv$bin_diff_mean[df_pred_bwd_inv$bin_mean == 6 & df_pred_bwd_inv$bin_diff_mean == -3] = 3

# Mode
# Make prediction direction invariant
# Dir 1
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 1 & df_pred_bwd_inv$bin_diff_mode == 4] = -2
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 1 & df_pred_bwd_inv$bin_diff_mode == 5] = -1
# Dir 2
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 2 & df_pred_bwd_inv$bin_diff_mode == 4] = -2
# Dir 4
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 4 & df_pred_bwd_inv$bin_diff_mode == -3] = 3
# Dir 5
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 5 & df_pred_bwd_inv$bin_diff_mode == -4] = 2
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 5 & df_pred_bwd_inv$bin_diff_mode == -3] = 3
# Dir 6
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 6 & df_pred_bwd_inv$bin_diff_mode == -5] = 1
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 6 & df_pred_bwd_inv$bin_diff_mode == -4] = 2
df_pred_bwd_inv$bin_diff_mode[df_pred_bwd_inv$bin_mode == 6 & df_pred_bwd_inv$bin_diff_mode == -3] = 3

```

- Get accuracy for target and 180° shifted from target from predictions
    - Average over folds
        - In the BWD case we do not have different folds. This is because the training data (betas) 
        was created while excluding BWD events in the GLM. In oder to keep the comparison clean
        we want to average over the four folds of the FWD data.
    - All further calculations are done based on the **mean** of angles within one TR

```{r, echo=FALSE}
# Calculate accuracy
# FWD
df_fwd_acc = ddply(df_pred_fwd_inv,
                   .(sub_id, mask, group, fold),
                   summarise,
                   accuracy_m120 = sum(bin_diff_mean == -2)/length(bin_diff_mean),
                   accuracy_m60 = sum(bin_diff_mean == -1)/length(bin_diff_mean),
                   accuracy_0 = sum(bin_diff_mean == 0)/length(bin_diff_mean),
                   accuracy_60 = sum(bin_diff_mean == 1)/length(bin_diff_mean),
                   accuracy_120 = sum(bin_diff_mean == 2)/length(bin_diff_mean),
                   accuracy_180 = sum(bin_diff_mean == 3)/length(bin_diff_mean))

# Average over folds
df_fwd_acc_subj = ddply(df_fwd_acc,
                        .(sub_id, mask, group),
                        summarise,
                        accuracy_m120 = mean(accuracy_m120),
                        accuracy_m60 = mean(accuracy_m60),
                        accuracy_0 = mean(accuracy_0),
                        accuracy_60 = mean(accuracy_60),
                        accuracy_120 = mean(accuracy_120),
                        accuracy_180 = mean(accuracy_180))

# BWD
df_bwd_acc = ddply(df_pred_bwd_inv,
                   .(sub_id, mask, group),
                   summarise,
                   accuracy_m120 = sum(bin_diff_mean == -2)/length(bin_diff_mean),
                   accuracy_m60 = sum(bin_diff_mean == -1)/length(bin_diff_mean),
                   accuracy_0 = sum(bin_diff_mean == 0)/length(bin_diff_mean),
                   accuracy_60 = sum(bin_diff_mean == 1)/length(bin_diff_mean),
                   accuracy_120 = sum(bin_diff_mean == 2)/length(bin_diff_mean),
                   accuracy_180 = sum(bin_diff_mean == 3)/length(bin_diff_mean))


```


- Fuse data frames of FWD and BWD events

```{r, echo=FALSE}
df_fwd_acc_subj$dir = 'fwd'
df_bwd_acc$dir = 'bwd'

df_acc_inv = rbind(df_fwd_acc_subj, df_bwd_acc)
```

---

# Calculate visual influence score

- Calculate difference between FWD 0° and BWD 0° classification accuracy in V1 and RSC
    - (BWD - FWD) accuracy
- Calculate difference between BWD 180° and FWD 180° accuracy in V1 and RSC
    - (FWD-BWD) accuracy (Turn around difference so the two values cancel each other out
    ==> No influence of absolute value)
    
    - Number above 0: BWD accuracy works better than FWD accuracy
    - Number below 0: FWD accuracy works better than BWD accuracy

```{r, echo=FALSE}
data_diff = ddply(df_acc_inv, 
                  .(sub_id, mask, group),
                  summarise,
                  diff_0 = accuracy_0[2] - accuracy_0[1] / accuracy_0[1],
                  diff_180 = accuracy_180[2] - accuracy_180[1] / accuracy_180[2])
```


- Calculate differences of these two differences for both masks:
    - $(\textrm{BWD}_{180} - \textrm{FWD}_{180}) - (\textrm{BWD}_{0} - \textrm{FWD}_{0})$
    - Value > 0: Visual influence (since effects don't cancel out), the more negative, the more visual 
    influence
    - Value < 0: No evidence for visual influence since prediction did not switch with BWD walking
    
```{r, echo=FALSE}
data_diff$score = data_diff$diff_180 - data_diff$diff_0
```


---

# Save visual infuence score data to .tsv

```{r}
file = file.path(repo_dir, 'analysis', 'backwards_walking', 'visual_influence.tsv', fsep=.Platform$file.sep)
write.table(data_diff, file, sep='\t', row.names = FALSE)
```

---

# Get BWD confusion function

```{r}
data_cm_bwd = data.frame(matrix(NA, 0, 8))
colnames(c(paste(seq(-120,180,by=60), 'deg', sep=''), 'sub_id', 'mask'))

# Loop over each mask
for(mask_count in c(1:length(masks))){
  
  # Select mask
  mask = masks[mask_count]
  
  for(sub_count in 1:length(sub_list)){
    
    # Loop over each participant
    sub_id = sub_list[sub_count]
    
    # Count number of occurences (levels assure you also count occurence of bins which did not occur)
    data = df_pred_bwd_inv[which(df_pred_bwd_inv$mask == mask & df_pred_bwd_inv$sub_id == sub_id),]
    counts = unname(table(factor(data$bin_diff_mean, levels=c(-2:3))))
    
    # Unroll unhandy table format of 'table' into data.frame
    counts = data.frame(matrix(as.numeric(as.character(counts)), 1, 6))
    colnames(counts) = c(paste(seq(-120,180,by=60), 'deg', sep=''))
    
    # Normalize counts on total number of predictions
    counts = counts/sum(counts)
    
    # Add mask and sub_id column
    counts$sub_id = sub_id
    counts$mask = mask
    
    # Append counts
    data_cm_bwd = rbind(data_cm_bwd, counts)
  }
}

# Add group column
data_cm_bwd$group = NA
data_cm_bwd$group[grep('old', data_cm_bwd$sub_id)] = 'older'
data_cm_bwd$group[grep('young', data_cm_bwd$sub_id)] = 'younger'

# Reorder columns
data_cm_bwd = data_cm_bwd[,c(7,8,9,1:6)]
```

- Save data frame of BWD CF to .tsv

```{r}
file = file.path(repo_dir, 'analysis', 'curve_fitting', 'data_cm_bwd.tsv', fsep=.Platform$file.sep)
write.table(data_cm_bwd, file, sep='\t', row.names = FALSE)
```

